---
title: "Question 2"
author: "Sina Sanei"
date: "October 30, 2018"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# a)
let's first import data and take a look at histogram: 



```{r data, echo=FALSE}
data=c(15.98247,16.47688,17.2935,19.29195,15.62418,19.17946,19.83762,17.66685,17.52842,14.2797,15.65064,15.46015,17.78624,16.52783,18.20742,16.9908,17.93268,21.80852,16.5106,18.25084,19.66132,15.68794,17.6264,15.05916,15.99123,16.53632,12.70026,16.52058,18.89823,16.33826,16.92829,17.40366,16.97416,15.52522,18.56712,17.70126,18.35067,14.88808,17.96294,16.6413,18.52277,17.60614,18.28284,17.21295,17.11902,18.32105,13.26564,16.90886,18.00639,17.81292,16.91042,18.82537,16.75111,15.87344,14.4237,14.79683,16.22904,17.07457,17.67224,16.62292,19.34927,16.12205,16.8359,16.30263,17.62287,15.9437,16.91413,18.1876,14.61401,18.94885,16.33257,18.65407,16.36635,16.30884,16.90533,19.1134,18.85181,16.55264,18.25012,20.19427,16.73713,17.90763,16.59451,16.27072,18.13678,15.63045,17.90084,15.0235,15.87718,15.21189,15.84529,14.22941,17.58531,18.95757,18.25935,18.36956,16.82062,16.63638,18.45567,17.42606)
hist(data, breaks = 10)
summary(data)
```

now lets define the log-likelihood function and first and second derivatives( for $\gamma = 0$ constatn) : 

```{r ll}
x = data
n = length(x)
density = function(x, mu){ 
  f = exp(-x+mu)/((1+exp(-x+mu))^2)
  return(f)}

loglik = function( x, mu,n){
  ll = mu*n - sum(x) - 2*sum(log(1+exp(-x+mu))) 
  return(ll)}


```
now begin newton-raphson algorithm : 

```{r nr}
x = data
mle_nr=function(xvec,stop_crit,sRate){
  startvalue=median(xvec);
  n=length(xvec);
  mu_curr=startvalue;
  nn=0
  mu_seq = NULL
  ####compute first derivative of log-likelihood #####
  first_derivll=n-2*sum(exp(-xvec+mu_curr)/(1+exp(-xvec+mu_curr)));
  ### Continue algorithm until the first derivative ###
  ### of the log-likelihood is within stop criterion ##
  while(abs(first_derivll)>stop_crit){
    ####compute second derivative of log-likelihood #####
    second_derivll=-2*sum(exp(-xvec+mu_curr)/(1+exp(-xvec+mu_curr))^2);
    #### Newton-Raphson’s  update of estimate of mu ####
    mu_new=mu_curr-sRate *(first_derivll/second_derivll);
    mu_seq = c(mu_seq, mu_new);
    mu_curr=mu_new;
    ####compute first derivative of log-likelihood #####
    first_derivll=n-2*sum(exp(-xvec+mu_curr)/(1+exp(-xvec+mu_curr)));
    nn=nn+1
  }
  return (list(thetahat=mu_curr, iterations =nn ,sequence = mu_seq))
}
``` 

```{r l}
## starting value = meadian(x)  , step size= 0.01 ,stop criterion =0.0001
a= mle_nr( x, 0.0001, 0.01)
plot(a$sequence)
a$iterations #number of iterations
a$thetahat
```

```{r l1}
## starting value = meadian(x) , step size= 0.01 ,stop criterion =0.001
a= mle_nr( x, 0.001, 0.01)
plot(a$sequence)
a$iterations #number of iterations
a$thetahat
```

```{r l2}
## starting value = meadian(x) , step size= 0.1 ,stop criterion =0.001
a= mle_nr( x, 0.001, 0.1)
plot(a$sequence)
a$iterations #number of iterations
a$thetahat
```

```{r l3}
## starting value = meadian(x) , step size= 1 ,stop criterion =0.001
a= mle_nr( x, 0.001, 1)
plot(a$sequence)
a$iterations #number of iterations
a$thetahat
```

for step size of larger than 1 algorithm diverges, 

# iii) 
using the observed fisher information as an estimate for expected fisher information , and pluging in the estimated $\mu$: 
```{r fi}
#second_derivll=-2*sum(exp(-xvec+mu_curr)/(1+exp(-xvec+mu_curr))^2)
second_derivll=-2*sum(exp(-x+17.09973)/(1+exp(-x+17.09973))^2)
```
hence the 95% confidence interval will be : 
```{r ci}
lb =17.09973 - qnorm(0.975)*1/ sqrt(-1*second_derivll)
ub=17.09973 + qnorm(0.975)*1/ sqrt(-1*second_derivll)
c(lb,ub)
```